---
title: "260_final"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "What Makes a Board Game Good?"
author: "Shimin Bi, Shi Chen, Yanying Qiu, Melissa Zhao"
date: "Dec 6, 2017"
output:
  html_document:
    toc: true
    toc_depth: 2
    df_print: paged
    number_sections: true
geometry: margin=1in
fontsize: 12pt
---
\newcommand{\noin}{\noindent}    
\newcommand{\Var}{\text{Var}}    
\newcommand{\Cov}{\text{Cov}}    
\newcommand{\bs}{\boldsymbol}
\newcommand{\mb}{\mathbf}

\allowdisplaybreaks

\definecolor{myblue}{RGB}{0,128,255}
\newenvironment{answer}{\par\color{myblue}}{\par}

# Overview and Motivation
## Background
We all love Settlers of Catan, but what is it about Catan that makes it so addictive? There are many different components of a board game that give it a lot of variations. Different board games can have various maximum or minimum number of players, varying length of play time, different themes, mechanisms, designers or even different difficulty! We are curious to find out which of these attributes actually make a board game a good one, measured by player ratings. With data collected from boardgamegeek.com, we have player ratings on thousands of board games and their features. We are interested to do some exploratory analysis on the features and potentially build models to predict which board games are more likely to be loved by players.

## Objectives
\begin{itemize}
  \item Investigate the possible traits of high-rating board games using ggplot2 package in R.
  \item Build models to predict the success of a new board game and find important features of the successful games (defined by high average player rating).
  \item Recommend board games to players based on certain specified criteria or other games that they love.
\end{itemize}

# Initial Questions
## Initial Questions
## New Questions 
## Further Thoughts

```{r, include=FALSE}
library(ggplot2)
library(dslabs)
library(tidyverse)
library(dplyr)
library(readr)
library(tidyr)
library(caret)
library(lazyeval)
library(randomForest)
library(proxy)
```

# Data
## Data Source
## Data Cleaning
### Replacing Wrong Values
We found several variables that need to be checked and replaced with appropriate values, including
\begin{itemize}
  \item min_players
  \item max_players
  \item weight
  \item avg_time
  \item min_time
  \item max_time
  \item year
\end{itemize}
Find all cells with value = 0 and cells with values not making sense. Search the original website to find information and overwrite them. Also recoding some continuos variables to categorical.
```{r}
game <- read.csv("bgg_data.csv")

### Overwriting non-sense cells
# min_players
game$min_players[1616] <- 1; game$min_players[1962] <- 1; game$min_players[2273] <- 1 
game$min_players[2408] <- 1; game$min_players[2693] <- 1; game$min_players[3332] <- 1 
game$min_players[3850] <- 1; game$min_players[3902] <- 1; game$min_players[4145] <- 1 
game$min_players[4247] <- 1; game$min_players[4311] <- 1; game$min_players[4442] <- 1 
game$min_players[4480] <- 1; game$min_players[4815] <- 1 

# max_players
game$max_players[1616] <- NA; game$max_players[1962] <- NA; game$max_players[2273] <- NA
game$max_players[2408] <- NA; game$max_players[2687] <- 2; game$max_players[2739] <- 4
game$max_players[2818] <- 2; game$max_players[3332] <- 1; game$max_players[3356] <- 4
game$max_players[3570] <- 2; game$max_players[3724] <- 2; game$max_players[3875] <- 2
game$max_players[3902] <- NA; game$max_players[4016] <- 2; game$max_players[4145] <- NA
game$max_players[4241] <- 2; game$max_players[4354] <- 2; game$max_players[4437] <- 2
game$max_players[4442] <- NA; game$max_players[4480] <- NA; game$max_players[4504] <- 2
game$max_players[4528] <- 2; game$max_players[4540] <- 2; game$max_players[4795] <- 2
game$max_players[4815] <- NA; game$max_players[4988] <- 2

# weight
game$weight[1477] <- NA; game$weight[4381] <- NA; game$weight[4521] <- NA

# min_time, avg_time and max_time are cleaned in excel and re-imported back
game <- read.csv("bgg_clean_dat.csv", sep = " ", header = T)
```

### Recoding
\begin{itemize}
  \item min_players - to categories single player, multi-player or party game
  \item max_players - to categories single player, multi-player or party game
  \item min_time - to categories 0 (short), 1(medium), 2(long)
  \item avg_time - to categories 0 (short), 1(medium), 2(long)
  \item weight - to categories 0(easy), 1(medium), 2(hard)
\end{itemize}
```{r}
# recode players
game$single_player = 0
game$single_player[game$min_players == 1] = 1
game$multi_player = 0
game$multi_player[game$min_players > 1 & game$max_players <= 4] = 1
game$party_player = 0
game$party_player[game$max_players > 4] = 1
# recode min_time
# 0 = short,  1 = medium, 2 = long
quantile(game$min_time, na.rm = T) 
game$cate_mintime = 0
game$cate_mintime[game$min_time >= 30 & game$min_time <= 90] = 1
game$cate_mintime[game$min_time > 90] = 2
# recode avg_time
# 0 = short,  1 = medium, 2 = long
quantile(game$avg_time, na.rm = T)
game$cate_avgtime = 0
game$cate_avgtime[game$avg_time >= 30 & game$avg_time <= 120] = 1
game$cate_avgtime[game$avg_time > 120] = 2
# recode weight
# 0 = easy, 1 = medium, 2 = hard
quantile(game$weight, na.rm = T)
game$cate_weight = 0
game$cate_weight[game$weight >= 1.73885 & game$weight <= 2.8889] = 1
game$cate_weight[game$weight > 2.8889] = 2

# write final cleaned csv and import csv
write.table(game, "bgg_final_clean_dat.csv", sep = "|")
game <- read.csv("bgg_final_clean_dat.csv", sep = "|", header = T)
```

### Hot-Coding
For each board game, it can have multiple mechanics or categories. We split the mechanics and categories into separate columns with each boradgame having 0 or 1 for each of the mechanic and category.
```{r}
# recode mechanic
# find unique mechanics
mech_str <- paste(as.character(game$mechanic), collapse = ", ")
mech_unique <- unique(strsplit(mech_str, ", ")[[1]])
mech_unique_lower <- unlist(lapply(mech_unique, function(x) {paste(strsplit(tolower(x), " ")[[1]], collapse = "_")}))

# create one empty column for each unique mechanic
mechanic_col <- data.frame(matrix(0, ncol = length(mech_unique_lower), nrow = dim(game)[1]))
colnames(mechanic_col) <- mech_unique_lower

# fill in the values of the mechanic columns
fill_mech_col <- function(df, mechanic_col) {
  for (i in 1:dim(df)[1]) {
    mech_col_num <- which(mech_unique %in% c(strsplit(as.character(df$mechanic[i]), ", ")[[1]]))
    for (j in mech_col_num) {
      mechanic_col[i, j] <- 1
    }
  }
  return(mechanic_col)
}
mechanic_col <- fill_mech_col(game, mechanic_col)

# recode categories
# find unique categories
cat_str <- paste(as.character(game$category), collapse = ", ")
cat_unique <- unique(strsplit(cat_str, ", ")[[1]])
cat_unique_lower <- unlist(lapply(cat_unique, function(x) {paste(strsplit(tolower(x), " ")[[1]], collapse = "_")}))

# create one empty column for each unique category
cat_col <- data.frame(matrix(0, ncol = length(cat_unique_lower), nrow = dim(game)[1]))
colnames(cat_col) <- cat_unique_lower

# fill in the values of the category columns
fill_cat_col <- function(df, cat_col) {
  for (i in 1:dim(df)[1]) {
    cat_col_num <- which(cat_unique %in% c(strsplit(as.character(df$category[i]), ", ")[[1]]))
    for (j in cat_col_num) {
      cat_col[i, j] <- 1
    }
  }
  return(cat_col)
}

cat_col <- fill_cat_col(game, cat_col)

df_new <- cbind(game, mechanic_col)
write.table(df_new, 'df_w_mechanic', sep = "|")
df_new2 <- cbind(game, cat_col)
write.table(df_new2, 'df_w_cat', sep = "|")

df_w_mechanic <- read.csv('df_w_mechanic', sep = "|")
df_w_cat <- read.csv('df_w_cat', sep = "|")
drops <- c('none')
df_mech_new <- df_w_mechanic[ , !(names(df_w_mechanic) %in% drops)]
df_mech_new$memory_mechanic <- df_mech_new$memory
df_mech_final <- df_mech_new[ , !(names(df_mech_new) %in% 'memory')]
df_cat_new <- df_w_cat[ , !(names(df_w_cat) %in% drops)]
df_cat_final <- df_cat_new[, 27:109]
df_recode_final <- cbind(df_mech_final, df_cat_final)
write.table(df_recode_final, 'df_recode_final_1127', sep = "|")
```


# Exploratory Data Analysis
```{r}
game <- read.csv("df_recode_final_1127", sep = "|")
```
## Summary Plots

### 1 The distribution of average rating and geek rating across years

```{r}

game %>% group_by(year) %>%
  filter(year > 1980 ) %>%
  summarise(avg_mean = mean(avg_rating), geek_mean = mean(geek_rating)) %>%
  ggplot()+
  geom_point(aes(year, avg_mean, col = 'avg_mean'))+
  geom_line(aes(year, avg_mean, col = 'avg_mean')) +
  geom_point(aes(year,  geek_mean, col = 'geek_mean'))+
  geom_line(aes(year,  geek_mean, col = 'geek_mean')) +
  scale_colour_manual(name="rating", values=c(avg_mean="red",geek_mean ="blue")) +
  ylab("rating") +
  xlab("year") +
  ggtitle('The distribution of average rating and geek rating across years') +
  theme_grey()
```


### 2 Changing difficulty of board games across years

```{r}

game %>% group_by(year) %>%
  filter(year > 1980) %>% 
  summarise(m = mean(weight, na.rm = T)) %>%
  ggplot(aes(year, m))+
  geom_point(color = 'blue')+
  geom_line(color = 'blue') +
  ylab("Difficulty") +
  xlab("year") +
  ggtitle('Changing difficulty of board games across years')+
  theme_grey()
```

### 3 The distribution of mean rank of board games across years

```{r}
game %>% group_by(year) %>%
  filter(year > 1980) %>%
  summarise(m = mean(rank)) %>%
  ggplot(aes(year, m))+
  geom_point(color = 'blue')+
  geom_line(color = 'blue') +
  ylab("Mean Rank") +
  xlab("year") +
  ggtitle('the distribution of mean rank of board games across years')+
  theme_grey()
```


##  4 Plots and ranking among 3 age groups
### 4.1 Generate agecat according IQR

```{r}
Q1 <- quantile(game$age, 0.25) 
Q3 <- quantile(game$age, 0.75) 
game <- game %>% mutate(agecat = ifelse (age %in% range(0,Q1), 1, ifelse(age %in% range(Q1, Q3), 2, 3)))
```

### 4.2 The top 10 rated (based on avg_rating) for each age group, game_id

```{r}
# agecat == 1 when age %in% range(0,Q1)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable
# agecat == 2 when age %in% range(Q1,Q3)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 2) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable
# agecat == 3 when age %in% range(Q3,)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 3) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable
```

### 4.3 The top 10 rated (based on avg_geek rating) for each age group, game_id

```{r}
# agecat == 1 when age %in% range(0,Q1)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable
# agecat == 2 when age %in% range(Q1,Q3)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 2) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable
# agecat == 3 when age %in% range(Q3,)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 3) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable
```

### 4.4 For each age group, get top 100 ranked games, and then find the top 10 rated (by freq)  mechanicsfor each age group.

```{r}
# agecat == 1 when age %in% range(0,Q1)
top100_1 <- game %>% 
  filter(agecat == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_1[,27:76]
m1 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
# agecat == 2 when age %in% range(Q1,Q3)
top100_2 <- game %>% 
  filter(agecat == 2) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_2[,27:76]
m2 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
# agecat == 3 when age %in% range(Q3,)
top100_3 <- game %>% 
  filter(agecat == 3) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_3[,27:76]
m3 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
```

### Plot for 3 age groups
### 4.5 Top 10 Ranked Game Mechanics for each Age Group

```{r}
library(RColorBrewer)
m11 <- data.frame(names=names(m1), m1)
m22 <- data.frame(names=names(m2), m2)
m33 <- data.frame(names=names(m3), m3)
p1 <- m11 %>% mutate(freq = m1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ffcccc") +
  theme_light() +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 0-8') +
  ylab('Frequency') +
  ylim(0,50)+
  coord_flip() 
p2 <- m22 %>% mutate(freq = m2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff9999") +
  theme_light() +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 9-12') +
  ylab('Frequency') +
  ylim(0,50)+
  coord_flip()
p3 <-  m33 %>% mutate(freq = m3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff4d4d") +
  theme_light() +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 13-21') +
  ylab('Frequency') +
  ylim(0,50)+
  coord_flip()
grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2),  ggplotGrob(p3),size = "last"))
```

### 4.6 Top 10 Ranked Game Categories for each Age Group

```{r}

c1 <- top100_1[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c2 <- top100_2[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c3 <- top100_3[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  

c11 <- data.frame(names=names(c1), c1)
c22 <- data.frame(names=c(names(c2)[1:7], 'manufacturing', names(c2)[9:10]), c2)
c33 <- data.frame(names=names(c3), c3)


pc1 <- c11 %>% mutate(freq = c1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ffcccc") +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 0-8') +
  ylab('Frequency') +
  ylim(0,40)+
  coord_flip() 
 

pc2 <- c22 %>% mutate(freq = c2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff9999") +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 9-12') +
  ylab('Frequency') +
  ylim(0,40)+
  coord_flip()

pc3 <-  c33 %>% mutate(freq = c3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff4d4d") +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 13-21') +
  ylab('Frequency') +
  ylim(0,40)+
  coord_flip()

grid.newpage()
grid.draw(rbind(ggplotGrob(pc1), ggplotGrob(pc2),  ggplotGrob(pc3),size = "last"))
```


## 5 Plots and ranking among 3 single_player, multi_player, party_player groups
### 5.1 Top 10 rated (based on avg_rating), game_id

```{r}
#single_player
game %>% 
  select(avg_rating, geek_rating, rank, single_player, game_id, names) %>%
  filter(single_player == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable

#multi_player
game %>% 
  select(avg_rating, geek_rating, rank, multi_player, game_id, names) %>%
  filter(multi_player == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable

#party_player
game %>% 
  select(avg_rating, geek_rating, rank, party_player, game_id, names) %>%
  filter(party_player == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable
```

### 5.2 The top 10 rated (based on avg_geek rating) for each player group, game_id

```{r}
#single_player
game %>% 
  select(avg_rating, geek_rating, rank, single_player, game_id, names) %>%
  filter(single_player == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable

#multi_player
game %>% 
  select(avg_rating, geek_rating, rank, multi_player, game_id, names) %>%
  filter(multi_player == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable

#party_player
game %>% 
  select(avg_rating, geek_rating, rank, party_player, game_id, names) %>%
  filter(party_player == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable
```

### 5.3 Plot the top 10 rated (avg_geek) for each player group, mechanic

```{r}
#single-player 
top100_1 <- game %>% 
   filter(single_player == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_1[,27:76]
m1 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
  
# multi-player 
top100_2 <- game %>% 
  filter(multi_player == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_2[,27:76]
m2 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  

# party-player
top100_3 <- game %>% 
  filter(party_player == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_3[,27:76]
m3 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  

# plot
m11 <- data.frame(names=names(m1), m1)
m22 <- data.frame(names=names(m2), m2)
m33 <- data.frame(names=names(m3), m3)


p1 <- m11 %>% mutate(freq = m1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ffcccc") +
  theme(axis.text=element_text(size=8)) +
  xlab('single_player') +
  ylab('Frequency') +
  ylim(0,50)+
  coord_flip() 
 

p2 <- m22 %>% mutate(freq = m2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff9999") +
  theme(axis.text=element_text(size=8)) +
  xlab('multi_player') +
  ylab('Frequency') +
  ylim(0,50)+
  coord_flip()

p3 <-  m33 %>% mutate(freq = m3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff4d4d") +
  theme(axis.text=element_text(size=8)) +
  xlab('party_player') +
  ylab('Frequency') +
  ylim(0,50)+
  coord_flip()

grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2),  ggplotGrob(p3),size = "last"))
```

### 5.4 Plot the top 10 rated (avg_geek) for each player group, categories

```{r}
c1 <- top100_1[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c2 <- top100_2[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c3 <- top100_3[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  

c11 <- data.frame(names=names(c1), c1)
c22 <- data.frame(names=names(c2), c2)
c33 <- data.frame(names=names(c3), c3)


pc1 <- c11 %>% mutate(freq = c1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ffcccc") +
  theme(axis.text=element_text(size=8)) +
  xlab('single_player') +
  ylab('Frequency') +
  ylim(0,40) +
  coord_flip() 
 

pc2 <- c22 %>% mutate(freq = c2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff9999") +
  theme(axis.text=element_text(size=8)) +
  xlab('multi_player') +
  ylab('Frequency') +
  ylim(0,40) +
  coord_flip()

pc3 <-  c33 %>% mutate(freq = c3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq), stat = "identity", fill= "#ff4d4d") +
  theme(axis.text=element_text(size=8)) +
  xlab('party_player') +
  ylab('Frequency') +
  ylim(0,40) +
  coord_flip()

grid.newpage()
grid.draw(rbind(ggplotGrob(pc1), ggplotGrob(pc2),  ggplotGrob(pc3),size = "last"))
```

# avg rating vs geek rating across game ranks by categories
```{r}
#loading game without separation of mech and cate into indicator variables 
game1 <- read.csv("bgg_final_clean_dat.csv", sep = "|", header = T)

library(purrr)
library(stringr)
# Functions
split_into_multiple <- function(column, pattern = ", ", into_prefix){
  cols <- str_split_fixed(column, pattern, n = Inf)
  cols[which(cols == "")] <- NA
  cols <- as_tibble(cols)
  m <- dim(cols)[2]
  names(cols) <- paste(into_prefix, 1:m, sep = "_")
  return(cols)
}

#Splitting 
game1 <- game1 %>% bind_cols(split_into_multiple(game1$category,',','category')) %>%
  bind_cols(split_into_multiple(game1$mechanic,',','mechanic'))

#Cleaning
game1 <- game1 %>% select(-category, -mechanic, -designer, -image_url)

#Tidying
tidygame <- game1 %>% gather(key, categories, category_1:category_11, na.rm = TRUE) %>% select(-key) %>%
  gather(key, mechanics, mechanic_1:mechanic_18, na.rm = TRUE) %>% select(-key)

tidygame$mechanics <- trimws(tidygame$mechanics)
tidygame$categories <- trimws(tidygame$categories)

#Categories vs ratings
tidygame %>% 
 group_by(categories) %>% summarize(avgrating = mean(avg_rating), avggeek = mean(geek_rating), avgrank = mean(rank)) %>%  
  ggplot() +
  geom_point(aes(reorder(categories, avgrank), avggeek, color = 'avg_geek'), size = 0.5) +
  geom_point(aes(categories, avgrating, color = 'avg_rating'), size = 0.5) + 
  scale_colour_manual(name="Rating", values=c(avg_geek="red", avg_rating="blue")) +
  theme(axis.text=element_text(size=8, angle = 60, hjust =1)) +
  ylab("Rating") +
  xlab("Rank") +
  ggtitle("Ratings by Categories") 
```

#avg rating vs geek rating across game ranks by mechanics
```{r}
#Mechanics vs ratings
tidygame %>% group_by(mechanics) %>% summarize(avgrating = mean(avg_rating), avggeek = mean(geek_rating), avgrank = mean(rank)) %>%
  ggplot() +
  geom_point(aes(reorder(mechanics, avgrank), avggeek, color = 'avg_geek'), size = 0.5) +
  geom_point(aes(mechanics, avgrating, color = 'avg_rating'), size = 0.5) +
  scale_colour_manual(name="Rating", values=c(avg_geek="red", avg_rating="blue")) + #adds legend
  theme(axis.text=element_text(size=9, angle = 60, hjust = 1)) +
  ylab("Rating") +
  xlab("Rank") +
  ggtitle("Ratings by Mechanics") 
```


# Beta Coefficients of Game Categories  (linear model with categories variable only)
```{r}

model1 <- lm(geek_rating ~ categories,data = tidygame)
coeff <- as.data.frame(model1$coefficients)
coeff$category <- as.factor(rownames(coeff))
colnames(coeff) <- c("coef","covariates")
confint_cat <- as.data.frame(confint(model1)) 

coeff_cat <- coeff[,]
coeff_cat$covariates <- gsub("categories", "", coeff_cat$covariates)
head(coeff_cat)

coeff_cat_error <- bind_cols(coeff_cat, confint_cat) 
colnames(coeff_cat_error) <- c("coef", "covariates", "lower", "upper")

ggplot(coeff_cat_error[-1,], aes( x=reorder(covariates, coef)))+
  geom_errorbar(aes(x =reorder(covariates, coef), ymin = lower, ymax = upper), color = "grey70") +
  geom_point(aes(y = coef), col = "#C1275C") +
  coord_flip() +
  labs(title = "Linear Model Coefficients for Category",y="Coefficients",x="Category")+
  scale_fill_gradient2(low = "light grey", mid = "grey70",
  high = "#C1275C", midpoint = 0.25) +
  theme_light() +
  guides(fill=guide_legend(title="Coefficient values")) +
  theme(axis.text=element_text(size=6))

```

# Beta Coefficients of Game Mechanics (linear model with mechanics variable only)
```{r}
model2 <- lm(geek_rating ~ mechanics, data = tidygame)
coeff <- as.data.frame(model2$coefficients)
coeff$category <- as.factor(rownames(coeff))
colnames(coeff) <- c("coef","covariates")
confint_mech <- as.data.frame(confint(model2)) 

coeff_mech <- coeff[,]
coeff_mech$covariates <- gsub("mechanics", "", coeff_mech$covariates)
coeff_mech_error <- bind_cols(coeff_mech, confint_mech) 
colnames(coeff_mech_error) <- c("coef", "covariates", "lower", "upper")

ggplot(coeff_mech_error[-1,], aes( x=reorder(covariates, coef)))+
  geom_errorbar(aes(x =reorder(covariates, coef), ymin = lower, ymax = upper), color = "grey70") +
  geom_point(aes(y = coef), col = "#C1275C") +
  coord_flip() +
  labs(title = "Linear Model Coefficients for Mechanics",y="Coefficients",x="Mechanics")+
  scale_fill_gradient2(low = "light grey", mid = "grey70",
  high = "#C1275C", midpoint = 0.25) +
  theme_light() +
  guides(fill=guide_legend(title="Coefficient values")) +
  theme(axis.text=element_text(size=8))
```

# Change in Rating Over the Years






# Change in Game Difficulty Over the Years






# Change in Game Popularity Over the Years 






# Change in Game Mechanics and Themes Over the Years
```{r}
# Change in game mechanics
df_new1 <- game %>% filter(year >= 1980)
df_new1$year_group <- cut(df_new1$year, breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010, 2015, 2018), include.lowest = TRUE)
dice_rolling <- df_new1 %>%
    group_by(year_group) %>%
    summarize(percent = sum(dice_rolling, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "dice_rolling")
hand_management <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(hand_management, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "hand_management")
variable_player_powers <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(variable_player_powers, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "variable_player_powers")
set_collection <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(set_collection, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "set_collection")
area_control_._area_influence <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(area_control_._area_influence, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "area_control_._area_influence")
card_drafting <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(card_drafting, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "card_drafting")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
mechanic_by_year <- as.data.frame(bind_rows(dice_rolling, hand_management, variable_player_powers, set_collection, area_control_._area_influence, card_drafting))
ggplot() + 
  geom_point(aes(x = dice_rolling$year_group, y = dice_rolling$percent, col = 'Dice Rolling')) + 
  geom_line(aes(x = dice_rolling$year_group, y = dice_rolling$percent, group = 1, col = 'Dice Rolling')) +
  geom_point(aes(x = hand_management$year_group, y = hand_management$percent, col = 'Hand Management')) + 
  geom_line(aes(x = hand_management$year_group, y = hand_management$percent, group = 1, col = 'Hand Management')) + 
  geom_point(aes(x = variable_player_powers$year_group, y = variable_player_powers$percent, col = 'Variable Player Powers')) +
  geom_line(aes(x = variable_player_powers$year_group, y = variable_player_powers$percent, group = 1, col = 'Variable Player Powers')) + 
  geom_point(aes(x = set_collection$year_group, y = set_collection$percent, col = 'Set Collection')) + 
  geom_line(aes(x = set_collection$year_group, y = set_collection$percent, group = 1, col = 'Set Collection')) +
  geom_point(aes(x = area_control_._area_influence$year_group, y = area_control_._area_influence$percent, col = 'Area Control/Area Influence')) + 
  geom_line(aes(x = `area_control_._area_influence`$year_group, y = area_control_._area_influence$percent, group = 1, col = 'Area Control/Area Influence')) +
  geom_point(aes(x = card_drafting$year_group, y = card_drafting$percent, col = 'Card Drafting')) + 
  geom_line(aes(x = card_drafting$year_group, y = card_drafting$percent, group = 1, col = 'Card Drafting')) +
  scale_colour_manual("", 
                      breaks = c("Dice Rolling", "Hand Management", "Variable Player Powers", "Set Collection", "Area Control/Area Influence", "Card Drafting"),
                      values = cbPalette[1:6]) +
  scale_x_discrete(breaks = dice_rolling$year_group, 
                   labels = seq(1980, 2015, 5)) +
  xlab("Year") +
  ylab("Percentage of games") +
  ggtitle("Evolution of Game Mechanics 1980 - 2018") +
  theme(legend.position="bottom")

# change in game categories
card_game <- df_new1 %>%
    group_by(year_group) %>%
    summarize(percent = sum(card_game, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "card_game")
wargame <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(wargame, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "wargame")
fantasy <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(fantasy, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "fantasy")
economic <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(economic, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "economic")
fighting <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(fighting, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "fighting")
science_fiction <- df_new1 %>%
    group_by(year_group) %>%
    summarise(percent = sum(science_fiction, na.rm = TRUE) / n()) %>%
    mutate(mechanic = "science_fiction")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")
ggplot() + 
  geom_point(aes(x = card_game$year_group, y = card_game$percent, col = 'Card Game')) + 
  geom_line(aes(x = card_game$year_group, y = card_game$percent, group = 1, col = 'Card Game')) +
  geom_point(aes(x = wargame$year_group, y = wargame$percent, col = 'War Game')) + 
  geom_line(aes(x = wargame$year_group, y = wargame$percent, group = 1, col = 'War Game')) + 
  geom_point(aes(x = fantasy$year_group, y = fantasy$percent, col = 'Fantasy')) +
  geom_line(aes(x = fantasy$year_group, y = fantasy$percent, group = 1, col = 'Fantasy')) + 
  geom_point(aes(x = economic$year_group, y = economic$percent, col = 'Economic')) + 
  geom_line(aes(x = economic$year_group, y = economic$percent, group = 1, col = 'Economic')) +
  geom_point(aes(x = fighting$year_group, y = fighting$percent, col = 'Fighting')) + 
  geom_line(aes(x = fighting$year_group, y = fighting$percent, group = 1, col = 'Fighting')) +
  geom_point(aes(x = science_fiction$year_group, y = science_fiction$percent, col = 'Science Fiction')) + 
  geom_line(aes(x = science_fiction$year_group, y = science_fiction$percent, group = 1, col = 'Science Fiction')) +
  scale_colour_manual("", 
                      breaks = c("Card Game", "War Game", "Fantasy", "Economic", "Fighting", "Science Fiction"),
                      values = cbPalette[1:6]) +
  scale_x_discrete(breaks = dice_rolling$year_group, 
                   labels = seq(1980, 2015, 5)) +
  xlab("Year") +
  ylab("Percentage of games") +
  ggtitle("Evolution of Game Categories 1980 - 2018") + 
  theme(legend.position="bottom")
```

# Predictions with Machine Learning
## Objective
We want to predict the success of a board game which is measured by its **average rating** on boardgamegeek.com.

## Methods
We tried four different machine learning methods, including linear regression, kth nearest neighbors, random forest, and support vector machine.

###Building train and test set
```{r}
library(caret)
set.seed(1)
game <- read.csv("df_recode_final_1127",  header =T, sep = "|") 

# drop the irrelevant columns like game ID, names, designer
# drop na
g1 <- game[, -c(2:4, 14, 16, 18:19, 77)] #4750  152
g <- drop_na(g1) # 4666  152

# Spliting data as training and test set. Using createDataPartition() function from caret
inTrain <- createDataPartition(y = g$avg_rating,
                               p=0.8)$Resample
train_set <- slice(g1, inTrain) 
test_set <- slice(g, -inTrain)

control <- trainControl(method = 'cv', number = 20)
```


### Linear Regression
```{r}
### Finding the best covariates to test with ML lm method 
model <- train(avg_rating ~ .,
             data = train_set,
             method = "lm",
             na.action=na.exclude,
             trControl = control,
             metric = "RMSE")
summary(model)

#Removing NAs
tidygame <- train_set %>% select(avg_rating, year, weight , single_player , multi_player , hand_management , point_to_point_movement , set_collection , variable_player_powers , card_drafting , area_control_._area_influence , campaign_._battle_card_driven , dice_rolling , simultaneous_action_selection , route.network_building , variable_phase_order , grid_movement , storytelling , worker_placement , deck_._pool_building , player_elimination , press_your_luck , hex.and.counter , stock_holding , betting.wagering , line_drawing , rock.paper.scissors , environmental , card_game , economic , wargame , fighting , city_building , farming , murder.mystery) %>% na.omit(.)

#Stepwise selection
lm.null <- lm(avg_rating ~ 1, data = train_set)
lm.full <- lm(avg_rating ~ year + weight + single_player + multi_player + hand_management + point_to_point_movement + set_collection + variable_player_powers + card_drafting + area_control_._area_influence + campaign_._battle_card_driven + dice_rolling + simultaneous_action_selection + route.network_building + variable_phase_order + grid_movement + storytelling + worker_placement + deck_._pool_building + player_elimination + press_your_luck + hex.and.counter + stock_holding + betting.wagering + line_drawing + rock.paper.scissors + environmental + card_game + economic + wargame + fighting + city_building + farming + murder.mystery, data = train_set)
mod1 <- step(lm.null, direction = "both", scope = list(lower = lm.null, upper = lm.full))

#finalized linear model 
mod1 <- lm(avg_rating ~ weight + year + wargame + single_player + 
    deck_._pool_building + storytelling + grid_movement + hex.and.counter + 
    campaign_._battle_card_driven + multi_player + city_building + 
    rock.paper.scissors + simultaneous_action_selection + area_control_._area_influence + 
    line_drawing + betting.wagering + route.network_building + 
    press_your_luck + murder.mystery + fighting + card_drafting + 
    stock_holding + variable_player_powers + card_game + point_to_point_movement, 
    data = train_set)
summary(mod1)

#predictions vs actual rating
predictions <- predict(mod1, test_set)
cor(predictions, test_set$avg_rating)

#graph of predictions vs actual rating
data.frame(test_set$avg_rating, predictions) %>% ggplot(aes(test_set$avg_rating, predictions)) +
  geom_point(color = "pink") + 
  stat_ellipse(color = "pink") +
  xlab("Actual Average Rating") +
  ylab("Predicted Average Rating") +
  ggtitle("Actual vs Predicted Average Rating")

#graph of linear model coefficients by magnitude
mod_coef <- data.frame(mod1$coeff) %>% mutate(variable = rownames(mod_coef))

ggplot(mod_coef[-1,], aes( x=reorder(variable, mod1.coeff), y=mod1.coeff, fill=mod1.coeff))+
  geom_bar(stat="identity") +
  coord_flip() +
  labs(title = "Final Linear Model Coefficients",y="Coefficients",x="Variable")+
  scale_fill_gradient(low="grey50", high="grey50")+
  theme_light() +
  guides(fill=guide_legend(title="Coefficient values"))
```

### K-Nearest Neighbors
```{r}
knnFit <- train(avg_rating ~.,
             data = train_set,
             method = "knn",
             na.action=na.exclude,
             trControl = control,
             preProcess = c("center", "scale"),
             tuneLength = 10)

knnFit

#plot it
knnFit %>% ggplot() +
  geom_point(color='pink') +
  geom_line(color ='pink')

#prediction
knnPredict <- predict(knnFit,newdata = test_set) # 932

#plot actual vs prediction
data <- data.frame(cbind(test_set$avg_rating,knnPredict))
names(data) <- c('avg_rating','knnPredict')

data %>% ggplot(aes(avg_rating, knnPredict)) +
  geom_point(color = 'pink') +
  xlim(c(6,8)) +
  ylim(c(6,8)) +
  ylab('KNN_prediction') +
  xlab('Actual average_rating')
```


### Random Forest
```{r}
# omit na
df1_na_omit <- na.omit(game)
# set seed
set.seed(1)

# create training and testing data
inTrain <- createDataPartition(y = df1_na_omit$geek_rating,
                               p=0.8)$Resample
train_set <- slice(df1_na_omit, inTrain) 
test_set <- slice(df1_na_omit, -inTrain)

# select needed columns
train_set_select <- train_set[, c(5:10, 11, 15, 20:110)]

# run randomForest with all features
fit <- randomForest(avg_rating ~  ., 
      data = train_set_select,
      ntree = 500)

# plot feature importance
impt <- as.data.frame(importance(fit))
impt$variable <- names(train_set_select[, -7])
impt <- transform(impt, variable = reorder(variable, IncNodePurity))
impt %>% 
  mutate(sort(IncNodePurity, decreasing = TRUE)) %>%
  ggplot() +
  geom_bar(aes(y = IncNodePurity, x = variable), stat = 'identity') + 
  coord_flip() +
  ylab("Feature Importance") + 
  xlab("Feature") +
  ggtitle("Board Game Features Ranked by Importance")
```

Top 20 Most Important Features
```{r}
# select top 20 most important features
impt_feature <- impt %>% 
  mutate(sort(IncNodePurity, decreasing = TRUE)) %>% 
  head(20) %>%
  select(variable)
impt_feature <- as.character(as.vector(impt_feature$variable))

impt_feature_df <- impt %>%
  filter(variable %in% impt_feature)
impt_feature_df %>% 
  mutate(sort(IncNodePurity, decreasing = TRUE)) %>%
  ggplot() +
  geom_bar(aes(y = IncNodePurity, x = variable), stat = 'identity') + 
  coord_flip() +
  ylab("Feature Importance") + 
  xlab("Feature") +
  ggtitle("Board Game Features Ranked by Importance")
```

```{r}
# fit the random forest model
# change number of variables randomly sampled as candidates at each split
RMSE_mtry <- c()
for (m in 1:50) {
  fit <- randomForest(avg_rating ~  ., 
      data = train_set_select,
      ntree = 100, 
      mtry = m)
  
  # make predictions on test set
  predictions <- predict(fit, test_set)

  # calculate RMSE
  RMSE <- sqrt(sum((predictions - test_set$avg_rating)^2)/length(predictions))
  print(RMSE)
  RMSE_mtry <- c(RMSE_mtry, RMSE)
}

which(RMSE_mtry == min(RMSE_mtry))

ggplot() +
  geom_line(aes(x = 1:50, y = RMSE_mtry)) + 
  geom_point(aes(x = 1:50, y = RMSE_mtry)) +
  ggtitle("Choose the Best Number of Variables to Include at Each Split") +
  xlab("mtry") +
  ylab("RMSE")

# change number of trees to grow
RMSE_ntree <- c()
for (n in 1:15) {
  fit <- randomForest(avg_rating ~  ., 
      data = train_set_select,
      ntree = n * 50, 
      mtry = 30)
  
  # make predictions on test set
  predictions <- predict(fit, test_set)

  # calculate RMSE
  RMSE <- sqrt(sum((predictions - test_set$avg_rating)^2)/length(predictions))
  print(RMSE)
  RMSE_ntree <- c(RMSE_ntree, RMSE)
}

ggplot() +
  geom_line(aes(x = seq(50, 750, 50), y = RMSE_ntree)) + 
  geom_point(aes(x = seq(50, 750, 50), y = RMSE_ntree)) +
  ggtitle("Choose the Best Number of Trees to Grow") +
  xlab("ntree") +
  ylab("RMSE")

which(RMSE_ntree == min(RMSE_ntree))
min(RMSE_ntree)
```

Fit a model using our best mtry and ntree.
```{r}
# best model
fit <- randomForest(avg_rating ~  ., 
      data = train_set_select,
      ntree = min(RMSE_ntree), 
      mtry = min(RMSE_mtry))

# make predictions
predictions <- predict(fit, test_set)

# calculate new RMSE
RMSE <- sqrt(sum((predictions - test_set$avg_rating)^2)/length(predictions))
print(RMSE)

# R^2
R_2 <- 1 - sum((test_set$avg_rating - predictions)^2) / sum((test_set$avg_rating - mean(test_set$avg_rating))^2)
R_2

# plot
ggplot() +
  geom_point(aes(x = test_set$avg_rating, y = predictions), col = "pink") + 
  ggtitle("Predictions vs True Average Ratings") +
  xlab("True Average Rating") +
  ylab("Predicted Average Rating")
```

### Support Vector Machines
```{r}
### support vector machine
g1 <- game[, -c(2:4, 14, 16, 18:19, 77)] #4750  152
g <- drop_na(g1) # 4666  152

# Spliting data into training and test set. 
inTrain <- createDataPartition(y = g$avg_rating, p=0.8)$Resample
train_set <- slice(g1, inTrain) 
test_set <- slice(g, -inTrain)

#control <- trainControl(method = 'cv', number = 20)
svmFit <- train(avg_rating ~.,
             data = train_set,
             method = "svmLinear",
             na.action=na.exclude,
             trControl = control,
             preProcess = c("center", "scale"),
             tuneLength = 10)
svmFit
svmPredict <- predict(svmFit,newdata = test_set) 

#plot actual vs prediction
data <- data.frame(cbind(svmPredict, test_set$avg_rating))
names(data) <- c('svmPredict','avg_rating')
data %>% ggplot(aes(avg_rating, svmPredict)) +
  geom_point(color =  'hotpink2') +
  xlim(c(6,8)) +
  ylim(c(6,8)) +
  xlab('Actual average_rating') +
  ylab('SVM_prediction')
```

# Board Game Recommender
We also built a simple board game recommender where a user can input their favorite board game and we will make recommendations of several board games that we think they might like, based on how similar the games are to their favorite game.
```{r}
# omit na
df1_na_omit <- na.omit(game)

# drop columns not wanted
drops <- c("rank", "bgg_url", "game_id", "image_url", "mechanic", "category", "designer")
df_rec <- df1_na_omit[ , !(names(df1_na_omit) %in% drops)]

# function to get similarity between two boardgames using Euclidean distance
get_most_simi <- function(game_name, df) {
  # get only the mechanics and category columns
  df_new <- df[, c(1, 20:153)]
  
  # create a vector of the features of the user's favorite game
  game_played <- as.numeric(as.vector(df_new[df_new$names == game_name, ]))[-1]
  
  # calculate Euclidean distance between user's favorite game and every other game in our data
  score <- numeric(0)
  for (i in 1:dim(df_new)[1]) {
    score <- c(score, 
               dist(list(game_played, as.numeric(df_new[i, -1])), method = "Euclidean"))
  }
  names(score) <- df_new[, 1]

  games <- names(score)
  score <- as.data.frame(score)
  score$game <- games
  score <- score[order(score$score), ]
  similar_games <- score %>% 
    filter(score < quantile(score, 0.02) & 
           score != 0)
  game_list <- df %>%
    filter(names %in% similar_games$game)
  recommendations <- game_list[order(game_list$geek_rating, decreasing = TRUE), ] %>%
    select(names) %>%
    head(10)
  return(recommendations)
}

# example
# input: Kingdom Builder
get_most_simi("Kingdom Builder", df_rec)
```

# Final Analysis