---
title: "260_project_11272017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dslabs)
library(tidyverse)
library(dplyr)
library(knitr)
library(readr)
ds_theme_set()

#game <- read.csv(file = file.choose()) #4999   20
#iqgame <- read.csv('bgg_db_2017_04.csv', header = TRUE)

# clean and recoded data

game <- read.csv("df_recode_final_1127",  header =T, sep = "|") 
names(game) #  4750  160
```

## 1 EDA - basic discription
### 1 ) missting value?
```{r}
missing(game) #none
```
 
### 2）Mechanic and category plot

```{r}
library(purrr)
library(stringr)
# Functions
split_into_multiple <- function(column, pattern = ", ", into_prefix){
  cols <- str_split_fixed(column, pattern, n = Inf)
  cols[which(cols == "")] <- NA
  cols <- as.tibble(cols)
  m <- dim(cols)[2]
  names(cols) <- paste(into_prefix, 1:m, sep = "_")
  return(cols)
}

#Splitting 
g <- game %>% bind_cols(split_into_multiple(game$category,',','category')) %>%
  bind_cols(split_into_multiple(game$mechanic,',','mechanic'))

#Cleaning
g <- g %>% select(-category, -mechanic, -designer, -image_url)

#Tidying
tidydata <- g %>% gather(key, categories, category_1:category_11, na.rm = TRUE) %>% select(-key) %>%
  gather(key, mechanics, mechanic_1:mechanic_18, na.rm = TRUE) %>% select(-key)

tidydata$mechanics <- trimws(tidydata$mechanics)
tidydata$categories <- trimws(tidydata$categories)


#Mechanics vs ratings
tidydata %>% group_by(mechanics) %>% summarize(avgrating = mean(avg_rating), avggeek = mean(geek_rating), avgrank = mean(rank)) %>%
  ggplot() +
  geom_point(aes(avggeek, reorder(mechanics, avgrank), color = 'avg_geek'), size = 0.5) +
  geom_point(aes(avgrating, mechanics, color = 'avg_rating'), size = 0.5) +
  scale_colour_manual(name="Rating", values=c(avg_geek="red", avg_rating="blue")) +
  theme(axis.text=element_text(size=4)) +
  ylab("Rank") +
  xlab("Rating") +
  ggtitle("Ratings by Mechanics") 

#Categories vs ratings
tidydata %>% 
 group_by(categories) %>% summarize(avgrating = mean(avg_rating), avggeek = mean(geek_rating), avgrank = mean(rank)) %>%  
  ggplot() +
  geom_point(aes(avggeek, reorder(categories, avgrank), color = 'avg_geek'), size = 0.5) +
  geom_point(aes(avgrating, categories, color = 'avg_rating'), size = 0.5) + 
  scale_colour_manual(name="Rating", values=c(avg_geek="red", avg_rating="blue")) +
  theme(axis.text=element_text(size=4)) +
  ylab("Rank") +
  xlab("Rating") +
  ggtitle("Ratings by Categories") 
```


### 3)  descriptive analysis of covariates (plots)


```{r}
game %>% ggplot(aes(geek_rating, avg_rating)) +
  geom_point(alpha = 0.8) +
   geom_smooth()
```

*the distribution of avg_rating across years & the distribution of geek_rating across years*

```{r}
game %>% group_by(year) %>%
  filter(year > 1980 ) %>%
  summarise(avg_mean = mean(avg_rating), geek_mean = mean(geek_rating)) %>%
  ggplot()+
  geom_point(aes(year, avg_mean, col = 'avg_mean'))+
  geom_line(aes(year, avg_mean, col = 'avg_mean')) +
  geom_point(aes(year,  geek_mean, col = 'geek_mean'))+
  geom_line(aes(year,  geek_mean, col = 'geek_mean')) +
  scale_colour_manual(name="rating", values=c(avg_mean="red",geek_mean ="blue")) +
  ylab("rating") +
  xlab("year") +
  ggtitle('The distribution of average rating and geek rating across years')
```


*the distribution of owned and votes across years*

```{r}
game %>% group_by(year) %>%
  filter(year > 1980) %>%
  summarise(sum2 = sum(owned)) %>%
  ggplot(aes(year, sum2))+
  geom_point(color ="blue")+
  geom_line(color ="blue") +
  ylab("total number of owned") +
  xlab("year") +
  ggtitle('The distribution of total owned across years')

# orginal  owned and votes across years*
game %>% group_by(year) %>%
  filter(year > 1980) %>%
  summarise(sum1 = sum(owned), sum2 = sum(num_votes)) %>%
  ggplot()+
  geom_point(aes(year, sum1, col = 'owned'))+
  geom_point(aes(year, sum2, col = 'votes'))+
  scale_colour_manual(name="total number of", values=c(owned="red", votes="blue")) +
  ylab("total number") +
  xlab("year") +
  ggtitle('The distribution of total owned and votes across years')

# only  4, game$owned == game$num_votes

```

*the distribution of weights  across years*

```{r}
game %>% group_by(year) %>%
  filter(year > 1980) %>% 
  summarise(m = mean(weight, na.rm = T)) %>%
  ggplot(aes(year, m))+
  geom_point(color = 'blue')+
  geom_line(color = 'blue') +
  ylab("Difficulty") +
  xlab("year") +
  ggtitle('Changing difficulty of board games across years')
```

*the distribution of ranks across years*

```{r}
game %>% group_by(year) %>%
  filter(year > 1980) %>%
  summarise(m = mean(rank)) %>%
  ggplot(aes(year, m))+
  geom_point(color = 'blue')+
  geom_line(color = 'blue') +
  ylab("Mean Rank") +
  xlab("year") +
  ggtitle('the distribution of mean rank of board games across years')
```


### 4) Age and age group

*Distribution of age*

```{r}
game %>% ggplot(aes(age)) +
  geom_histogram()
```

*generate 3 age groups based on interquartile*

#### (1) generate agecat according IQR*

```{r}
Q1 <- quantile(game$age, 0.25) 
Q3 <- quantile(game$age, 0.75) 
game <- game %>% mutate(agecat = ifelse (age %in% range(0,Q1), 1, ifelse(age %in% range(Q1, Q3), 2, 3)))
```


#### (2).1 top 10 rated (avg_rating) for each age group, game_id*

```{r}
# agecat == 1 when age %in% range(0,Q1)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable

# agecat == 2 when age %in% range(Q1,Q3)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 2) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable

# agecat == 3 when age %in% range(Q3,)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 3) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable
```


 *top 10 rated (avg_geek) for each age group, game_id*
 
```{r}
# agecat == 1 when age %in% range(0,Q1)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable

# agecat == 2 when age %in% range(Q1,Q3)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 2) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable

# agecat == 3 when age %in% range(Q3,)
game %>% 
  select(avg_rating, geek_rating, rank, age, agecat, game_id, names) %>%
  filter(agecat == 3) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable
```

#### (2).2 for each age group, get top 100 ranked games, and then find the ctop 10 rated (freq) for each age group, mechanics*
mechanic 27 -76 

```{r}
# agecat == 1 when age %in% range(0,Q1)
top100_1 <- game %>% 
  filter(agecat == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_1[,27:76]
m1 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
  
# agecat == 2 when age %in% range(Q1,Q3)
top100_2 <- game %>% 
  filter(agecat == 2) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_2[,27:76]
m2 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  

# agecat == 3 when age %in% range(Q3,)
top100_3 <- game %>% 
  filter(agecat == 3) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_3[,27:76]
m3 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
```


*Plot for 3 age groups*

```{r}
library(RColorBrewer)
m11 <- data.frame(names=names(m1), m1)
m22 <- data.frame(names=names(m2), m2)
m33 <- data.frame(names=names(m3), m3)



p1 <- m11 %>% mutate(freq = m1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low ="yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 0-8') +
  ylab('Frequency') +
  coord_flip() 
 

p2 <- m22 %>% mutate(freq = m2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low ="yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 9-12') +
  ylab('Frequency') +
  coord_flip()

p3 <-  m33 %>% mutate(freq = m3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low ="yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 13-21') +
  ylab('Frequency') +
  coord_flip()


require("gridExtra")
grid.arrange(arrangeGrob(p1, p2, p3))
```


#### (2).3 top 10 rated for each age group, categories*
category 78-last


```{r}
c1 <- top100_1[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c2 <- top100_2[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c3 <- top100_3[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  

c11 <- data.frame(names=names(c1), c1)
c22 <- data.frame(names=c(names(c2)[1:7], 'manufacturing', names(c2)[9:10]), c2)
c33 <- data.frame(names=names(c3), c3)


pc1 <- c11 %>% mutate(freq = c1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low ="yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 0-8') +
  ylab('Frequency') +
  coord_flip() 
 

pc2 <- c22 %>% mutate(freq = c2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low ="yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 9-12') +
  ylab('Frequency') +
  coord_flip()

pc3 <-  c33 %>% mutate(freq = c3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low ="yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('Age 13-21') +
  ylab('Frequency') +
  coord_flip()

grid.arrange(arrangeGrob(pc1, pc2, pc3))
```


#### (3).1 Table for top 10 rated for each player

*top 10 rated (avg_rating), game_id*

```{r}
#single_player
game %>% 
  select(avg_rating, geek_rating, rank, single_player, game_id, names) %>%
  filter(single_player == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable

#multi_player
game %>% 
  select(avg_rating, geek_rating, rank, multi_player, game_id, names) %>%
  filter(multi_player == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable

#party_player
game %>% 
  select(avg_rating, geek_rating, rank, party_player, game_id, names) %>%
  filter(party_player == 1) %>% 
  arrange(desc(avg_rating)) %>% 
  head(10) %>% kable
```

*top 10 rated (avg_geek), game_id*

```{r}
#single_player
game %>% 
  select(avg_rating, geek_rating, rank, single_player, game_id, names) %>%
  filter(single_player == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable

#multi_player
game %>% 
  select(avg_rating, geek_rating, rank, multi_player, game_id, names) %>%
  filter(multi_player == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable

#party_player
game %>% 
  select(avg_rating, geek_rating, rank, party_player, game_id, names) %>%
  filter(party_player == 1) %>% 
  arrange(desc(geek_rating)) %>% 
  head(10) %>% kable
```


#### (3).2 top 10 rated (avg_geek) for each player group, mechanics*

```{r}

#single-player 
top100_1 <- game %>% 
   filter(single_player == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_1[,27:76]
m1 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  
  
# multi-player 
top100_2 <- game %>% 
  filter(multi_player == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_2[,27:76]
m2 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  

# party-player
top100_3 <- game %>% 
  filter(party_player == 1) %>% 
  arrange(rank) %>%
  mutate(new_rank = 1:n()) %>%
  filter(new_rank <= 100) 
mechanic <- top100_3[,27:76]
m3 <- mechanic %>% colSums() %>% sort(decreasing = T) %>% head(10)  

# plot
m11 <- data.frame(names=names(m1), m1)
m22 <- data.frame(names=names(m2), m2)
m33 <- data.frame(names=names(m3), m3)


p1 <- m11 %>% mutate(freq = m1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low = "yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('single_player') +
  ylab('Frequency') +
  coord_flip() 
 

p2 <- m22 %>% mutate(freq = m2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low = "yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('multi_player') +
  ylab('Frequency') +
  coord_flip()

p3 <-  m33 %>% mutate(freq = m3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low = "yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('party_player') +
  ylab('Frequency') +
  coord_flip()

grid.arrange(arrangeGrob(p1, p2, p3))
```


#### (3).3 top 10 rated (avg_geek) for each player group, categories*

```{r}
c1 <- top100_1[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c2 <- top100_2[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  
c3 <- top100_3[,78:160] %>% colSums() %>% sort(decreasing = T) %>% head(10)  

c11 <- data.frame(names=names(c1), c1)
c22 <- data.frame(names=names(c2), c2)
c33 <- data.frame(names=names(c3), c3)


pc1 <- c11 %>% mutate(freq = c1) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low = "yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('single_player') +
  ylab('Frequency') +
  coord_flip() 
 

pc2 <- c22 %>% mutate(freq = c2) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low = "yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('multi_player') +
  ylab('Frequency') +
  coord_flip()

pc3 <-  c33 %>% mutate(freq = c3) %>%
  ggplot () +
  geom_bar(aes(x = reorder(names, freq), y= freq,  fill= freq), stat = "identity") +
  scale_fill_gradient(low = "yellow", high = 'purple') +
  theme(axis.text=element_text(size=8)) +
  xlab('party_player') +
  ylab('Frequency') +
  coord_flip()

grid.arrange(arrangeGrob(pc1, pc2, pc3))
```



# Modeling
## Try knn (Chen), svm (Shimin, Chen) Cross validation to choose parameters
      
# KNN
```{r}
library(caret)
set.seed(1)
game <- read.csv("df_recode_final_1127",  header =T, sep = "|") 

# drop the irrelevant columns like game ID, names, designer
# drop na
g1 <- game[, -c(2:4, 14, 16, 18:19, 77)] #4750  152
g <- drop_na(g1) # 4666  152

# Spliting data as training and test set. Using createDataPartition() function from caret
inTrain <- createDataPartition(y = g$avg_rating,
                               p=0.8)$Resample
train_set <- slice(g1, inTrain) 
test_set <- slice(g, -inTrain)

control <- trainControl(method = 'cv', number = 20)
knnFit <- train(avg_rating ~.,
             data = train_set,
             method = "knn",
             na.action=na.exclude,
             trControl = control,
             preProcess = c("center", "scale"),
             tuneLength = 10)

knnFit

```


#plot it
```{r}
plot(knnFit)
```


#prediction
```{r}
knnPredict <- predict(knnFit,newdata = test_set) # 932
length(knnPredict)
length(test_set$avg_rating)
```

#plot actual vs prediction

```{r}
data <- data.frame(cbind(knnPredict,  test_set$avg_rating))
names(data) <- c('knnPredict','avg_rating')

data %>% ggplot(aes(knnPredict, avg_rating)) +
  geom_point(color =  'blue') +
  xlim(c(6,8)) +
  ylim(c(6,8)) +
  xlab('KNN_prediction') +
  ylab('Actual average_rating')

```





```{r}
#Plotting yields Number of Neighbours Vs RMSE (based on repeated cross validation)
plot(knnFit, print.thres = 0.5, type="S")
```



```{r}
library(pROC)
knnPredict <- predict(knnFit,newdata = test_set )
knnROC <- roc(test_set$avg_rating,knnPredict)
knnROC
```


```{r}
data <- data.frame(cbind(knnPredict,  test_set$avg_rating))
names(data) <- c('knnPredict','avg_rating')

data %>% ggplot(aes(knnPredict, avg_rating)) +
  geom_point(color =  'blue') +
  xlim(c(6,8)) +
  ylim(c(6,8)) 

#+
 # stat_ellipse(lwd=2, type='norm')

```



